The imprint of global magma oceans on exoplanet demographics

### 2023-01-06 src/tex/ms.tex
new figure: stat. power grid in x_H2O, f_rgh

\section{Results}
\subsection{Statistical signature of the runaway greenhouse threshold}
To characterize the population-level imprint of individual radius changes, we generated a generic planet population with an injected runaway greenhouse effect assuming a water fraction $x_\mathrm{H_2O} = \wrr$.
\begin{figure}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/HnullHmo.pdf}
        \caption{Planets above and below the runaway greenhouse threshold.
        \textit{Top}: Planet state as a function of orbital period. Planets with and without a runaway greenhouse climate mix and are not distinguishable in period space.
        \textit{Bottom}: Radii of synthetic planets with injected radius deviation as a function of instellation. Only the planets marked as transiting are observable. Above the runaway greenhouse threshold $S_\mathrm{thresh} =\SI{280}{\watt\per\meter\squared} $, the radii of a fraction of planets are inflated by the amount indicated with gray lines. The sharp boundary at $S_\mathrm{thresh}$ causes a discontinuity in the average planet radius (blue). This runaway greenhouse hypothesis can be tested against the null hypothesis $H_0$ (green), where average radii are independent of instellation.}
        \label{fig:HnullHmo}
    \end{centering}
\end{figure}
Figure~\ref{fig:HnullHmo} shows the resulting planetary radii.
When ordered in orbital period space, the different planet types overlap, diluting the demographic imprint.
With net instellation as an independent variable, planets above and below the runaway greenhouse threshold separate.
We also show the predictions of observable average planet radii from the statistical hypotheses defined above.
%The value for the average radius \textit{outside} the runaway greenhouse regime was computed from our nominal observed sample, whereas the radius change in the magma ocean case compared to the null hypothesis is a prediction from the magma ocean model, evaluated on the same sample.
Within the runaway greenhouse limit, an average radius change of \avgRadiusChange\ occurs.

@@

>>>>

\section{Results}
\subsection{Statistical signature of the runaway greenhouse threshold}
To characterize the population-level imprint of individual radius changes, we generated a generic planet population with an injected runaway greenhouse effect assuming a water fraction $x_\mathrm{H_2O} = \wrr$.
\begin{figure}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/HnullHmo.pdf}
        \caption{Synthetic planets above and below the runaway greenhouse threshold.
        \textit{Top}: Planet state as a function of orbital period. Planets with and without a runaway greenhouse climate mix and are not distinguishable in period space.
        \textit{Bottom}: Radii of synthetic planets with injected radius deviation as a function of instellation. Only the planets marked as transiting are observable. Above the runaway greenhouse threshold $S_\mathrm{thresh} =\SI{280}{\watt\per\meter\squared} $, the radii of a fraction of planets are inflated by the amount indicated with gray lines. The sharp boundary at $S_\mathrm{thresh}$ causes a discontinuity in the average planet radius (blue). This runaway greenhouse hypothesis can be tested against the null hypothesis $H_0$ (green), where average radii are independent of instellation.}
        \label{fig:HnullHmo}
    \end{centering}
\end{figure}
Figure~\ref{fig:HnullHmo} shows the resulting planetary radii.
When ordered in orbital period space, the different planet types overlap, diluting the demographic imprint.
With net instellation as an independent variable, planets above and below the runaway greenhouse threshold separate.
We also show the predictions of observable average planet radii from the statistical hypotheses defined above.
%The value for the average radius \textit{outside} the runaway greenhouse regime was computed from our nominal observed sample, whereas the radius change in the magma ocean case compared to the null hypothesis is a prediction from the magma ocean model, evaluated on the same sample.
Within the runaway greenhouse limit, an average radius change of \avgRadiusChange\ occurs.

@@

\begin{note}
    Given an optimistic survey design, what influences the probability of correctly rejecting a false null hypothesis?
    \begin{figure}[ht!]
%    \script{figurescript.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/optimistic_statpwr_H2O-f.pdf}
            \caption{
                Statistical power of the runaway greenhouse hypothesis test as a function of model parameters.
                For a sample size $N = \Nplanets$, the color code shows the fraction of simulations resulting in a sound detection ($\Delta \ln \mathcal{Z} > 3$) for different combinations of water mass fraction and fraction of greenhouse climate planets.
                Higher values in either parameter result in a more reliable detection.
            }
            \label{fig:statpwr_H2O-f}
        \end{centering}
    \end{figure}
    Figure~\ref{fig:statpwr_H2O-f} explores the statistical power achieved in the above scenario for different combinations of the poorly constrained parameters $x_{H_2O}$ and $f_\mathrm{rgh}$.
    The higher the water content and thus the radius change, and the higher the fraction of planets in runaway greenhouse climates, the greater the prospect of a statistically sound detection.
    As long as this fraction is larger than $\sim 0.1$, a sample size of \Nplanets\ is sufficient for a $\SI{50}{\percent}$ detection rate even for water ratios as low as $x_{H_2O} = 0.001$.
\end{note}
